---
title: "Homework 4"
author: "Juliette J. Rubin"
date: "March 5, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library("raster")
library("rgeos")
library("rgdal")
library("sp")
library("betareg")

```

## Bats on the landscape

I have chosen to study tree bat roost sites, with an environmental predictor of moonlight gradiant across a landscape

a) While some bats roost and hunt in open space, many bats are lunarphobic and thus will only fly in dark, canopy-protected areas. Additionally, tree bats roost singly in tree hollows or under tree bark, rather than in group-oriented caves.

b) The moonlight predictor is highly spatially autocorrelated (low phi) because tropical secondary forest is usually dense and patchy, and thus the darker areas on a landscape will be clumped and the moonlit, open areas will also be somewhat continuous


```{r bats}

rmvn <- function(n, mu = 0, V = matrix(1)) {
  p <- length(mu)
  if (any(is.na(match(dim(V), p)))) 
    stop("Dimension problem!")
  D <- chol(V)
  t(matrix(rnorm(n * p), ncol = p) %*% D + rep(mu, rep(n, p)))
}
#you can increase SD by adding ,sd=10 after (n*p). In the given rnorm function the sd is given at 1, so you have to specify if you want it to be different.
#making matrix from draws from random dist

# Set up a square lattice region
simgrid <- expand.grid(1:50, 1:50)
n <- nrow(simgrid)

# Set up distance matrix
distance <- as.matrix(dist(simgrid))

#larger phi is less autocorrelated
phi = .04 
plot(1:100, exp(-phi * 1:100), type = "l", xlab = "Distance", ylab = "Correlation")

X <- rpois(n, lambda=exp(-1+rmvn(1, rep(0, n), exp(-phi * distance))))


#X<-rbinom(n, prob=plogis(-5+rmvn(1,rep(0,n),exp(-phi*distance))),size=3)



# Visualize results
Xraster <- rasterFromXYZ(cbind(simgrid[, 1:2] - 0.5, X))

plot(Xraster)

#Converting raster to a dataframe
spat_dat=rasterToPoints(Xraster)
head(spat_dat)
nrow(spat_dat)

#pulling out environmental covariate from sample points
GO = sample(x=c(1:nrow(spat_dat)),size=40)

points(spat_dat[GO,c(1:2)])

#creating parameters for bat presence amidst moonlight
presence_intercept=-2.7
presence_slope=0.8

#zero inflated model/hurdle model
PA <- rbinom(40,plogis(presence_intercept+spat_dat[GO,3]*presence_slope),
             size=1)


#Determining bat abundance

count_intercept=-1.9
count_slope=0.8
over_dispersion=0.5



#PA is part of zero inflated process - if we draw 0 then abundance doesn't matter because present/absent data indicates that it isn't there
abundance=PA*rnbinom(40,mu=
                       exp(count_intercept+count_slope*spat_dat[GO,3]),
                     size=over_dispersion)

plot(abundance~spat_dat[GO,3])

#to give to partner
write.csv(data.frame(abundance,spat_dat[GO,3]),file="fake_data.csv")
```


c) sample organism points are constrained by the number of radiotelemetry units we can deploy on the landscape (40)

e) we might observe a hurdle model because in localities of intense moonlight, there will likely be no bats, while in dark forested areas we are interested in the abundance of individuals (since we know they will be there)

f) We assume that these distributions are positive and that they are overdispersed


```{r ptarmigan}

#working with the data frame from Clara
icybirds<-read.csv("icybirds.csv")

#coding presence/absence
PA1=ifelse(icybirds$abundance>0,1,0)

#creating a binomial model with the data just to measure presence/absence as a function of %cover of ice
glm(PA1~icybirds$environmental_covariate,family="binomial")
#intercept: -2.14, slope: 0.827
plot(PA1~icybirds$environmental_covariate)


#subsetting data by presence
abundance2 <- icybirds[-which(icybirds$abundance==0),]
abundance2

#running a glm with this subsetted data to find coefficients to use in beta distribution
glm(abundance2$abundance~abundance2$environmental_covariate, family="poisson")
#intercept: 0.851, slope:0.866

ice_cover=runif(100,min=0,max=1)
#recapturing coefficients by first simulating from the beta distribution
mean_relationship=plogis(0.851+0.866*ice_cover^2)
fractional_cover<-rbeta(100,shape1=mean_relationship*.5,shape2=1-mean_relationship*.5)
plot(fractional_cover~icybirds$environmental_covariate)

#building the betareg model
icybeta_model<-betareg(fractional_cover~icybirds$environmental_covariate)
coef(icybeta_model)
#intercept: -0.398, environmental_covariate(%ice cover): -0.02, phi:0.78
# These data are quite overdispersed because small phi
plogis(-0.398)
#baseline ptarmigan per 30m^2 landsat pixel is 0.40

```

i) we can measure autocorrelation by assessing phi, where a larger phi indicates less autocorrelation and a smaller phi indicates more.
